---
title: "EDS241: Assignment 1"
author: "Paloma Cartwright"
date: '`r Sys.Date()`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3,
                      echo = TRUE, message = FALSE, warning = FALSE)

# load packages
packages <- c("stargazer", "here", "tidyr", "tidyverse",
           "dplyr", "stringr", "janitor", "huxtable",
           "cowplot", "ggplot2", "tinytex", "datasets",
           "tibble", "readxl", "estimatr", "car")

for (i in packages) {
  if (require(i, character.only = TRUE) == FALSE) {
    install.packages(i, repos = "http://cran.us.r-project.org")
  }
  else {
    require(i, character.only = TRUE)
  }
}


```

\noindent The data for this assignment come from CalEnviroScreen 4.0, a mapping and data tool produced by the California Office of Environmental Health Hazards Assessment (OEHHA). The data are compiled and constructed from a variety of sources and cover all 8,035 census tracts in California. Source: https://oehha.ca.gov/calenviroscreen/report/calenviroscreen-40 .

\noindent The full data are contained in the file CES4.xls, which is available on Gauchospace (note that the Excel file has three “tabs” or “sheets”). The data is in the tab “CES4.0FINAL_results” and “Data Dictionary” contains the definition of the variables.

# Clean and plot data

\noindent The following code loads and cleans the data.

```{r , include=TRUE}

# Load data and clean names
ces4_raw <- read_excel("CES4.xlsx", sheet = "CES4.0FINAL_results") %>%
  na_if("NA") %>%
  clean_names()


# Select columns wanted data
ces4 <- ces4_raw %>%
  select("census_tract", "total_population",
         "california_county", "low_birth_weight", "pm2_5", "poverty")

ces4_clean <- na.omit(ces4)

```

\noindent The following code chunks answer the questions provided in the assignment.  

(a)
```{r , include=TRUE}
# Avg PM2.5 Conc

avg_pm25_conc <- mean(ces4$pm2_5)

```

The average PM2.5 concentration across all census tracts in California is `r avg_pm25_conc`.

(b) 
```{r}
# highest poverty

highest_poverty <- ces4 %>%
  group_by(california_county) %>%
  summarize(avg_poverty = na.omit(mean(poverty))) %>%
  arrange(desc(avg_poverty))

```

The county with the highest level of poverty is: . 

(c) The histograms are shown below:

```{r}

ggplot(data = ces4_clean, aes(x = low_birth_weight)) +
  geom_bar()

ggplot(data = ces4_clean, aes(x = pm2_5)) +
  geom_histogram()
```
(d) 
```{r}
model <- lm_robust(low_birth_weight ~ pm2_5, data = ces4)
summary(model)
```

The slope coefficient for the regression of Low Birth Rate on PM2.5 is `r model$coefficients[2]`. The heteroskedacticity-robust standard error is `r model$std.error[2]`. The slope coefficient tells us that for every 1 microgram per cubic meter increase in PM2.5 concentration, there is a `r model$coefficients[2]` increase in the percentage of census tract births with weight less than 2500g. This is statistically significant at the 5% level. 

(e)
```{r}

```


(f) 
```{r}
model2 <- lm_robust(low_birth_weight ~ pm2_5 + poverty, data = ces4)
summary(model2)
```

The estimated coefficient for poverty is `r model2$coefficients[3]`. This tells us that holding PM2.5 constant, for every one percent increase in the population of the census tract living below twice the federal poverty line there is a `r model2$coefficients[3]` increase in the percentage of census tract births with weight less than 2500g. The coefficient for PM2.5 increases which displays that when poverty is considered in the regression, the PM2.5 value becomes more prominent. These variables are correlated. In the model that omits poverty, the regression coefficient of PM2.5 captures the effect of poverty on low birth weights.  




(g)
```{r}
linhyp <- linearHypothesis(model2, c("pm2_5 = 0", "poverty=0"),
                           white.adjust = "hc2")
linhyp
```
With the p-value being `r linhyp$'Pr(>Chisq)'[2]`, we can reject the null hypothesis that the effect of PM2.5 is equal to the effect of poverty. 

