---
title: "EDS241: Assignment 1"
author: "Paloma Cartwright"
date: '`r Sys.Date()`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)

# load packages
packages=c("stargazer", "here", "tidyr", "tidyverse", "dplyr","stringr", "janitor", "huxtable",
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "readxl", "estimatr", "patchwork", "car")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}


```

\noindent The data for this assignment come from CalEnviroScreen 4.0, a mapping and data tool produced by the California Office of Environmental Health Hazards Assessment (OEHHA). The data are compiled and constructed from a variety of sources and cover all 8,035 census tracts in California. Source: https://oehha.ca.gov/calenviroscreen/report/calenviroscreen-40 .

\noindent The full data are contained in the file CES4.xls, which is available on Gauchospace (note that the Excel file has three “tabs” or “sheets”). The data is in the tab “CES4.0FINAL_results” and “Data Dictionary” contains the definition of the variables.

# Clean and plot data

\noindent The following code loads and cleans the data.

```{r , include=TRUE}

# Load data and clean names
CES4_raw <- read_csv(here("ces4.csv")) %>% 
  clean_names()


# Select columns wanted data
ces4 <- CES4_raw %>%
  select(census_tract, total_population, 
         california_county, low_birth_weight, pm2_5, poverty)

```

\noindent The following code chunks answer the questions provided in the assignment.  

# (a) What is the average concentration of PM2.5 across all census tracts in California?

```{r , include=TRUE}
# Avg PM2.5 Conc

avg_pm25_conc <- mean(ces4$pm2_5)

```

The average PM2.5 concentration across all census tracts in California is `r avg_pm25_conc`.

# (b) What county has the highest level of poverty in California? 

```{r}
# highest poverty 

highest_poverty <- ces4 %>% 
  group_by(california_county) %>% 
  summarize(avg_poverty = na.omit(mean(poverty))) %>% 
  arrange(desc(avg_poverty))

```

The county with the highest level of poverty is: . 

# (c) Make a histogram depicting the distribution of percent low birth weight and PM2.5.

```{r}

birth_weight_plot <- ggplot(data = ces4) +
  geom_histogram(aes(x = low_birth_weight)) + 
  theme_minimal()+
  labs(y = "Frequency", 
       x = "% of Birth Weight < 2500 grams")

pm2_5_plot <- ggplot(data = ces4) +
  geom_histogram(aes(x = pm2_5)) +
  theme_minimal() +
  labs(x = "Frequency", 
       y = "PM 2.5 in micrograms per cubic meter")

```

```{r, fig.cap="California Census Track Distributions of Low Birth Weight and PM2.5"}
birth_weight_plot | pm2_5_plot
```


# (d) Estimate a OLS regression of `low_birth_weight` on `pm2_5`. Report the estimated slope coefficient and its heteroskedasticity-robust standard error. Interpret the estimated slope coefficient. Is the effect of `pm2_5` on `low_birth_weight` statistically significant at the 5%?

```{r}
model <- lm_robust(low_birth_weight ~ pm2_5, data = ces4)
summary(model)
```

The slope coefficient for the regression of Low Birth Weight on PM2.5 is `r model$coefficients[2]`. The heteroskedacticity-robust standard error is `r model$std.error[2]`. The slope coefficient tells us that for every 1 microgram per cubic meter increase in PM2.5 concentration, there is a `r model$coefficients[2]` increase in the percentage of census tract births with weight less than 2500g. This is statistically significant at the 5% level. 

# (e) Suppose a new air quality policy is expected to reduce PM2.5 concentration by 2 micrograms per cubic meters. Predict the new average value of `low_birth_weight` and derive its 95% confidence interval. Interpret the 95% confidence interval.

```{r}
intercept <- model$coefficients[1]
b1 <- model$coefficients[2]
x = -2

```


# (f) Add the variable `poverty` as an explanatory variable to the regression in (d). Interpret the estimated coefficient on Poverty. What happens to the estimated coefficient on PM25, compared to the regression in (d). Explain. 

```{r}
model2 <- lm_robust(low_birth_weight ~ pm2_5 + poverty, data = ces4)
summary(model2)
```

The estimated coefficient for poverty is `r model2$coefficients[3]`. This tells us that holding PM2.5 constant, for every one percent increase in the population of the census tract living below twice the federal poverty line there is a `r model2$coefficients[3]` increase in the percentage of census tract births with weight less than 2500g. The coefficient for PM2.5 increases which displays that when poverty is considered in the regression, the PM2.5 value becomes more prominent. These variables are correlated. In the model that omits poverty, the regression coefficient of PM2.5 captures the effect of poverty on low birth weights. 


# (g) From the regression in (f), test the null hypothesis that the effect of PM2.5 is equal to the effect of Poverty

```{r}
linhyp <- linearHypothesis(model2, c("pm2_5 = 0", "poverty=0"),
                           white.adjust = "hc2")
linhyp
```
With the p-value being `r linhyp$'Pr(>Chisq)'[2]`, we can reject the null hypothesis that the effect of PM2.5 is equal to the effect of poverty. 

